---
title: "Data Mining - HW 1"
date: "Thursday, January 15, 2015"
output: html_document
authors: Tzu-Yuan Lin, Marcus A. Streips
---
```{r message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(lattice)
library(rpart)
library(party)
# library(partykit)
library(vcd)
library(vcdExtra)
library(arules)
library(plyr)
library(dplyr)
library(C50)
library(AppliedPredictiveModeling)
library(caret)
library(e1071)

data <- read.csv("salary-class.csv")
```
authors: "Tzu-Yuan Lin, Marcus A. Streips"

**Question 1**
  
**(a) Read the data set. Our goal is to predict INCOME. How many records do we have?**

The data set has **32,561** records. Using the R `str` command gives us:

'data.frame':        32561 obs. of  11 variables:


**(b) What are the types of each field be (interval, ordinal, nominal, flag)?**

- AGE     : int   **(interval/continuous)**
- EMPLOYER: Factor w/ 9 levels " ?"," Federal-gov",..: **(nominal)**
- DEGREE  : Factor w/ 16 levels " 10th"," 11th",..:  **(ordinal)**
- MSTATUS : Factor w/ 7 levels " Divorced"," Married-AF-spouse",..: **(nominal)**
- JOBTYPE : Factor w/ 15 levels " ?"," Adm-clerical",..:  **(nominal)**
- SEX     : Factor w/ 2 levels " Female"," Male": . **(flag)**
- C.GAIN  : int  2174 0 0 0 0 0 0 0 14084 5178 ... **(interval/continuous)**
- C.LOSS  : int  0 0 0 0 0 0 0 0 0 0 ... **(interval/continuous)**
- HOURS   : int  40 13 40 40 40 40 16 45 50 40 ... **(interval/continuous)**
- COUNTRY : Factor w/ 42 levels " ?"," Cambodia",..: .**(nominal)**
- INCOME  : Factor w/ 2 levels " <=50K"," >50K": **(flag)**
  
**(c) What effect does each of these three fields seem to have on INCOME?**

```{r boxplot, fig.height=4, fig.width=4,echo=FALSE}
boxplot(AGE ~ INCOME, data, xlab="Income Interval", ylab="Age")
freqtbl_age <-as.data.frame((with(data, table(INCOME, AGE))))
ggplot(freqtbl_age, aes(INCOME, AGE)) +
        geom_tile(aes(fill = Freq), colour = "black") +
        scale_fill_gradient(low = "white", high = "steelblue")
freqtbl_sex <- as.data.frame((with(data, table(INCOME, SEX))))
ggplot(freqtbl_sex, aes(INCOME, SEX)) +
        geom_tile(aes(fill = Freq), colour = "black") +
        scale_fill_gradient(low = "white", high = "steelblue")
freqtbl_degree <-as.data.frame((with(data, table(INCOME, DEGREE))))
ggplot(freqtbl_degree, aes(INCOME, DEGREE)) +
        geom_tile(aes(fill = Freq), colour = "black") +
        scale_fill_gradient(low = "white", high = "steelblue")

```
  
As seen from the figures above, those **earning >=$50 K** are expected to **live longer** on average, be **male**, and are **less** likely to have dropped out of school before 12th grade. 
  
**Question 2**
  
**(a) Find the support and confidence for the rule "if department = systems then salary = 46K..50K"**
  
Using the `apriori` command from the `arules` package we get the following:  
  
```{r, eval=FALSE}
   lhs                     rhs                      support confidence     lift

12 {department=systems} => {salary=46K..50K}      0.1388889  0.7142857 1.978022
```
  

**(b) Which of these three sets of rules has the lowest misclassification rate?**
  
Using the `C5.0` command from the `C50` package we determine that **AGE** has the lowest misclassification rate at **27.8%** (see results below), when compared to status (50%) and department (52.8%)(results not shown).  
  
This is confirmed by the SPSS analysis attached. 

```{r, eval=FALSE}
Evaluation on training data (180 cases):

                Rules     
	  ----------------
	    No      Errors

	     5   50(27.8%)   <<


	   (a)   (b)   (c)   (d)   (e)   (f)    <-classified as
	  ----  ----  ----  ----  ----  ----
	    50                                  (a): class 26K..30K
	          40                            (b): class 31K..35K
	                 5                      (c): class 36K..40K
	          10                            (d): class 41K..45K
	     5    30                30          (e): class 46K..50K
	           5                       5    (f): class 66K..70K
```
